---
title: "Assignment 1 for Coursera Reproducible" 
output: html_document
---
# Loading Data

``` {r}

activity <- read.csv("activity.csv")
str(activity)

```


```{r}

activity$date <- as.Date(activity$date)
str(activity)

```

Count Missing NAs


```{r}

sum(is.na(activity$steps))

```

## Total Steps Per Day?

Need to subset the data to create clean set.

```{r}

activity_new<-activity[which(!is.na(activity$steps)),]

```


```{r}

average <-tapply(activity_new$steps, activity_new$date, sum)

```

## Create a Histogram

Create a simple histogram using base plotting method.

```{r}

hist(average,10, main = "Aggregate Steps Per Day", xlab = "")

```

### Calculate Mean and Median

```{r}
mean(average)
```

```{r}
median(average)
```


## What is the average daily activity pattern?

```{r}

daily <- tapply(activity_new$steps, activity_new$interval, mean)
plot(y = daily, x = names(daily), type = "l", xlab = "5-Minute-Intervals", 
    main = "Daily Activity Pattern", ylab = "Average Steps")

```


Find the highest interval:

```{r}


daily[daily==max(daily)]

```


# Inputing missing values

Count NAs and Fill NAs with medians of intervals. 

```{r}
sum(is.na(activity))
```

### Fill! 

```{r}

activity_new <- activity
activity_new[which(is.na(activity_new$steps)),1]<-
        daily[as.character(activity_new[which(is.na(activity_new$steps)),3])]

```

Check for mistakes:

```{r}

sum(is.na(activity_new))

```


## Create Histogram

```{r, fig.height=10, fig.width=10}


perday_new <- tapply(activity_new$steps, activity_new$date, sum)
```




``` {r}
par(mfrow=c(1,2))
hist(perday_new,10, main = "Total number of steps taken per day", xlab = "Steps"
     , ylim =c(0, 25))
abline(v = median(perday_new), col = 4, lwd = 4)
hist(perday_new,10, main = "Total number of steps taken per day  
     (missing values replaced with mean of interval)", xlab = "Steps",
     ylim =c(0, 25))
abline(v = median(perday_new), col = 4, lwd = 4)

```


### We now calculate the median and the mean of the filled in dataset

```{r}
mean(perday_new)
```

```{r}

median(perday_new)
```

The differences is minimal for the two different data sets:

*Mean*
Unfilled: 10766
Filled: 10766

*Median*
Unfilled:10766
filled: 10766

# Are there differences in activity patterns between weekdays and weekends?

Filter between the weekend and weekday:

```{r, fig.height=20, fig.width=20}

activity_new$wd<-weekdays(activity_new$date)
activity_new$fwd<- as.factor(c("weekend", "weekday"))
activity_new[activity_new$wd == "Sunday" | activity_new$wd == "Saturday" ,5]<- factor("weekend")
activity_new[!(activity_new$wd == "Sunday" | activity_new$wd == "Saturday"),5 ]<- factor("weekday")

```


```{r, fig.height=20, fig.width=20}

activity_new_we <- subset(activity_new, fwd == "weekend") 
activity_new_wd <- subset(activity_new, fwd == "weekday") 
dailyactivity_we <- tapply(activity_new_we$steps, activity_new_we$interval, mean)
dailyactivity_wd <- tapply(activity_new_wd$steps, activity_new_wd$interval, mean)
par(mfrow=c(2,1))
plot(y = dailyactivity_wd, x = names(dailyactivity_wd), type = "l", xlab = "5-Minute Interval", 
     main = "Daily Activity Pattern on Weekdays", ylab = "Average number of steps", 
     ylim =c(0, 250))
plot(y = dailyactivity_we, x = names(dailyactivity_we), type = "l", xlab = "5-Minute Interval", 
     main = "Daily Activity Pattern on Weekends", ylab = "Average number of steps", 
     ylim =c(0, 250))
```

The differences between weekday and week show that on the weekend, the activity is more sustained. This is different than the week, where there is a definite period of minimal activity. 
